Improving the communication and transparency of stock assessment using interactive

### 2018-11-15 manuscript.Rmd
Rough draft of modeling section intro; set-up sugue to NCAM flexdashboard

In the age of data-intensive science and big data, scientists across many fields are faced with the challenge of synthesizing and communicating information from large and complex data sets. Scientific data are often presented as tabular summaries and static plots, however, seeing patterns in tables can be difficult and static plots preclude detailed explorations of the underlying data. New open source tools are bridging this gap by making it easier to build interactive figures [@perkel2018]. In particular, browser-based data visualizations have been playing an increasing prominent role in communicating information on a wide range of topics. Interactive graphics are frequently utilized by media outlets, such as the FiveThirtyEight, and scientists from a wide range of disciplines are starting to apply interactive visualization tools to explore and communicate their results [@yeatman2018; @tushar2017; @wick2015]. This trend has been fueled by the development of JavaScript libraries (e.g. D^3^ [@bostock2011], plotly [@plotly], leaflet [@leaflet], highcharts [@highcharts]) that enable the production of dynamic and interactive data visualizations in modern web-browsers. Interfaces to JavaScript libraries have also been developed to allow interactive plots to be coded using commonly used languages such as R [@R] or Python [@python]. By laying the technical infrastructure for creating interactive graphics, these new tools are ushering in a new era of Big Data by facilitating the exploration and communication of complex data and models.

@@

The data sets used in stock assessments are constantly growing. This growth in data either stems from the continuation of long-term monitoring efforts or from the addition of new monitoring programs. As such, stock assessment biologists often have to manage large volumes of data from a variety of sources. For instance, time series of reported landings and catch-at-age are "fishery-dependent" data that are frequently used in stock assessments. These data are often analyzed in conjunction with data from "fishery-independent" surveys that track changes in abundance and, in many cases, also monitor trends in biological factors such as age composition, growth rates, sex ratios and maturation stages. For some data-rich stocks, mark and recapture studies are also carried out to estimate movement, migration, growth rate, natural mortality, and discard mortality. All of the above-mentioned data sets are multidimensional and as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and discover key patterns within each. This challenge can be mediated, to a degree, by the application of interactive visualization tools as that they allow the construction of visual representations of the data that demonstrates large-scale patterns while also allowing the detailed exploration of the underlying data. For instance, the ability to zoom in on features or areas of interest and hover over specific points to reveal more information creates an interactive user-driven experience that expedites explorations of the data. This is exemplified by a recently developed interactive mapping tool developed for the exploration of a long-term tagging study.

@@

Synthesizing data from multiple sources presents a key challenge to stock assessment. Analyses of different data sources were traditionally carried out independently and the summaries or parameters from these analyses were used in the assessment model. This approach, however, is less than ideal because information may be lost and uncertainty may be unaccounted for when we "do statistics on the statistics" [@link1999; @maunder2013]. Such issues have largely been curtailed in contemporary stock assessments thanks to advances in software that have facilitated the analysis of all available data, in as raw a form as appropriate, in a single integrated analysis [@maunder2013]. Specifically, the application of statistical modeling tools such as JAGS [@plummer2003], AD Model Builder [@fournier2012] and Template Model Builder [@kristensen2015] allow the construction of a joint likelihood for an array of observations to, in theory, extract as much information as possible about the biological and fishery processes. However, integrated analyses are not a panacea because model misspecifications and data conflicts are an inevitable consequence of simplifying reality to a small series of equations [@maunder2017]. A potential solution to this quandary to use a superensemble model, whereby multiple models with different structures are run and their predictions are supplied as covariates to an additional superensemble model [@anderson2017]. Ensemble approaches reduce the risk of picking the "wrong" model and also expands the range of hypotheses explored [@dietterich2000]. These advances greatly improve our ability to assess the status and trends of fish populations, however, modern stock assessment biologists are now faced with the overwhelming task of understanding an ever expanding array of data inputs and model outputs.

- *TODO: tweak this intro paragraph and work up sugue to NCAM and the dashboard*

>>>>

In the age of data-intensive science and big data, scientists across many fields are faced with the challenge of synthesizing and communicating information from large and complex data sets. Scientific data are often presented as tabular summaries and static plots, however, seeing patterns in tables can be difficult and static plots preclude detailed explorations of the underlying data. New open source tools are bridging this gap by making it easier to build interactive figures [@perkel2018]. In particular, browser-based data visualizations have been playing an increasingly prominent role in communicating information on a wide range of topics. Interactive graphics are frequently utilized by media outlets, such as the FiveThirtyEight, and scientists from a wide range of disciplines are starting to apply interactive visualization tools to explore and communicate their results [@yeatman2018; @tushar2017; @wick2015]. This trend has been fueled by the development of JavaScript libraries (e.g. D^3^ [@bostock2011], plotly [@plotly], leaflet [@leaflet], highcharts [@highcharts]) that enable the production of dynamic and interactive data visualizations in modern web-browsers. Interfaces to JavaScript libraries have also been developed to allow interactive plots to be coded using commonly used languages such as R [@R] or Python [@python]. By laying the technical infrastructure for creating interactive graphics, these new tools are ushering in a new era of Big Data by facilitating the exploration and communication of complex data and models.

@@

The data sets used in stock assessments are constantly growing. This growth in data either stems from the continuation of long-term monitoring efforts or from the addition of new monitoring programs. As such, stock assessment biologists often have to manage large volumes of data from a variety of sources. For instance, time series of reported landings and catch-at-age are "fishery-dependent" data that are frequently used in stock assessments. These data are often analyzed in conjunction with data from "fishery-independent" surveys that track changes in abundance and, in many cases, also monitor trends in biological factors such as age composition, growth rates, sex ratios and maturation stages. For some data-rich stocks, mark and recapture studies are also carried out to estimate movement, migration, growth rate, natural mortality, and discard mortality. All of the above-mentioned data sets are multidimensional and as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and discover key patterns within each. This challenge can be mediated, to a degree, by the application of interactive visualization tools as that they allow detailed exploration of the data behind the plot. For instance, the ability to zoom in on features or areas of interest and hover over specific points to reveal more information creates an interactive user-driven experience that expedites explorations of the data. This is exemplified by an interactive mapping tool developed for the exploration of a long-term tagging study.

@@

Synthesizing data from multiple sources presents a key challenge to stock assessment. Analyses of different data sources were traditionally carried out independently and the summaries or parameters from these analyses were used in the assessment model. This approach, however, is less than ideal because information may be lost and uncertainty may be unaccounted for when we "do statistics on the statistics" [@link1999; @maunder2013]. Such issues have largely been curtailed in contemporary stock assessments thanks to advances in software that have facilitated the analysis of all available data, in as raw a form as appropriate, in a single integrated analysis [@maunder2013]. Specifically, statistical modeling tools such as JAGS [@plummer2003], AD Model Builder [@fournier2012] and Template Model Builder [@kristensen2015] allow the construction of a joint likelihood for an array of observations to, in theory, extract as much information as possible about the biological and fishery processes. However, integrated analyses are not a panacea because model misspecifications and data conflicts are an inevitable consequence of simplifying reality to a small series of equations [@maunder2017]. A potential solution to this quandary to use a superensemble model, whereby multiple models with different structures are run and their predictions are supplied as covariates to an additional superensemble model [@anderson2017]. Ensemble approaches reduce the risk of picking the "wrong" model and also expands the range of hypotheses explored [@dietterich2000]. These advances greatly improve our ability to assess the status and trends of fish populations, however, modern stock assessment biologists are now faced with the overwhelming task of understanding an ever expanding array of data inputs and model outputs. This challenge is compounded by the ever increasing demand for rapid and sound fisheries advice. Here we argue that the process of providing fisheries advice can be further streamlined by utilizing dynamic and interactive visualization tools to explore, diagnose and communicate stock assessment models. We hope to demonstrate this using a recently developed interactive tool for exploring an integrated assessment model for Northern cod.

# Notes

- Note that these tools facilitate the peer-review process and also improves the transparency and reproducibility of assessment models
- Dynamic document can be self contained
